---
title: "New York Education System Analysis"
author: "Krishnachandran Menon, Rohit Deshmukh, Shivam Thakur"
date: "2022-10-14"
output:
  
  html_document: default
  pdf_document: default
---
                              
                              
                              
                              
                           

### **Contents**
**1.	Problem Statement**

**2.  About the dataset** 

**3.	Output**  

   3.1 Institution Analysis in 2021
   
   3.2 Comparative Analysis
   
**4.	Conclusion** 

**5.	References**  


### **1. Problem Statement**

New York State Education Department is an entity that looks over the supervision of all public schools in New York. They are also the entity that is responsible for ensuring standardizing testing as well as the production and administration of state tests and Regents Examinations.

GRAD_RATE_AND_OUTCOMES_2021 gives us information regarding the graduation and dropout data for the state of New York, as well as for the counties necessary to interpret certain downfalls within the system in specific locations post analyzing the correlations between the economically disabled, students suffering from a disability, non-English speaking students and migrant student vs the graduation percentage and the drop rate to deduce necessary inferences and take actions for continuous improvement in the year of 2022 based on performance index of 2021.

The analysis conducted are:
- Enrollment Comparison by County
- School Comparison by highest enrollment
- County Average by Education Score
- English vs Non-English Speakers enrollment in different cohorts on a statewide level
- Migrant enrollments in different counties in the years 2020 & 2021 
- Race Wise Comparisons – Dropout Rate v Graduation Rate
- Comparison of graduation and dropout rate for Economically v Non-Economically challenged sections of the population
- Parameter comparison of Disabled students and All students
- Male v Female enrollment comparison done across the state of New York


### **2. About the dataset**

GRAD_RATE_AND_OUTCOMES_2021: 
The variables available for analysis are:
-  Aggregation_Index
- Aggregation_Type
- Aggregation_Name
- Membership_Desc
- Subgroup_Name
- Enroll_Cnt
- Grad_Cnt
- Grad_Pct
- Dropout_Cnt
- Dropout_Pct
- ged_pct
- non-diploma_credential_pct
- regent_pct
- regent_adv_pct
- local_pct

This database contains annual graduation, and dropout data for the state as well as by county, Need to Resource Capacity group, district, public school, and charter school. Annual graduation data is included for the current four-year cohort (June and August graduates), five-year June and August, and six-year June and August cohorts. 

Referred from Graduation Rate Database:  (https://data.nysed.gov/downloads.php)
 

### **3. Output**
#### **3.1 Institution Analysis in 2021**



```{r School comparison, warning=FALSE, include=FALSE}

library(sf)
library(raster)
library(dplyr)
library(magrittr)
library(tidyr)
library(tidyverse)
library(dplyr, warn.conflicts = FALSE)
library(spData)
library(ggplot2)
library(plotly)
countys <- read.csv("GRAD_RATE_AND_OUTCOMES_2021.csv")
```

#### **Enrollment Comparison by County**

The number of enrollments helps us confirm the population index of New York Demographic and ascertain the level of education of the populous in the year 2021.

```{r echo=FALSE, warning=FALSE}
#Total number of Registrations in all counties
options(dplyr.summarise.inform = FALSE)
countys <- read.csv("GRAD_RATE_AND_OUTCOMES_2021.csv") %>% 
filter(aggregation_type == "School") %>% 
dplyr::select(county_name,aggregation_name, enroll_cnt)  %>% 
group_by (aggregation_name, county_name) %>% 
summarise(Total_Enrollments=sum(enroll_cnt)) %>%  
arrange(desc(Total_Enrollments))  




 
county_school<- aggregate(Total_Enrollments ~ county_name, data=countys, sum) #using aggregate function to sum the values

county_school_final <- county_school[order(-county_school$Total_Enrollments),] #Ordering according to total enrollments

Top7_counties<-head(county_school_final,7) #Selecting top 7 to visualize in a better
options(scipen = 999)  #to remove return the actual value on the plot


fig<-ggplot(Top7_counties, aes(x=county_name, y=Total_Enrollments)) +
  geom_segment( aes(x=county_name, xend=county_name, y=0, yend=Total_Enrollments)) +
  geom_point( size=5, color="#0091D5", fill=alpha("#1C4E80", 0.3), alpha=0.7, shape=21, stroke=2)


fig+labs(title="Enrollment comparison by County",
        x ="County", y = "Enrollments")+theme(plot.title = element_text(hjust=0.5))
```

Post analysis it can be observed that the enrollment is high in Kings County accompanied by Queens County, Suffolk County, and Nassau County among others; matching the New York population index by county inferencing, the counties having the highest population are among the counties having highest enrollment too.


#### **School Comparison by highest enrollment**

A deeper analysis is conducted to confirm the above inference; In this analysis, the enrollments are computed with respect to schools in the entire state. A regular bar plot is plotted to understand the comparison.  

```{r Schools with highest enrollments in the state, echo=FALSE, warning=FALSE}

options(dplyr.summarise.inform = FALSE)

countys <- read.csv("GRAD_RATE_AND_OUTCOMES_2021.csv") %>% 
filter(aggregation_type == "School") %>% 
dplyr::select(county_name,aggregation_name, enroll_cnt)  %>% 
group_by (aggregation_name, county_name) %>% 
summarise(Total_Enrollments=sum(enroll_cnt)) %>%  
arrange(desc(Total_Enrollments))  
 
top7<-head(countys,7)

fig <- plot_ly(top7, x = ~aggregation_name, y = ~Total_Enrollments, type = 'bar',marker = list(color ='#1C4E80'))
fig <- fig %>% layout(title = "Top 10 Schools with highest number of Enrollments in NYS (Year 2020-21) ",
         xaxis = list(title = "Schools"),
         yaxis = list(title = "Enrollments")) %>% 
         layout(xaxis = list(categoryorder = "total descending"))
fig
```

The plot shows Brentwood High School, Suffolk having the highest enrollment in the year 2021	followed by Brooklyn Technical High School, Kings, Fort Hamilton High School, Kings, and Francis Lewis High School, Queens. The counties having the highest population consist of the schools having the highest population hence bolstering the support for the inference.




#### **County Average by Education Score**

A metric called Education Score is introduced to find the best-performing counties. It is defined as 80% of mean graduation percentage – 20% of mean dropout percentage across all cohorts. The top 10 counties are shown below. 


```{r  echo=FALSE, warning=FALSE}
#Average Graduation Rate and Drop Rate by County
df_grad_2021 <- read.csv('GRAD_RATE_AND_OUTCOMES_2021.csv')
df_grad_2020 <- read.csv('GRAD_RATE_AND_OUTCOMES_2020.csv')

#Average Graduation Rate and Drop Rate by County
df <- df_grad_2021 %>% 
filter(df_grad_2021$aggregation_type=='County')%>%
dplyr::select(subgroup_name,grad_cnt, dropout_cnt, enroll_cnt, county_name) %>%
filter(subgroup_name=='All Students') %>%
group_by(county_name) 

df$grad_cnt <- as.numeric(df$grad_cnt)
df$dropout_cnt <- as.numeric(df$dropout_cnt)
df$enroll_cnt <- as.numeric(df$enroll_cnt)

df_county_mean_grad <- aggregate(grad_cnt ~ county_name, data=df, mean)
df_county_mean_dropout <- aggregate(dropout_cnt ~ county_name, data=df, mean)
df_county_mean_enroll <- aggregate(enroll_cnt ~ county_name, data=df, mean)

df_county_mean_enroll$enroll_cnt <- as.numeric(unlist(df_county_mean_enroll$enroll_cnt))
df_county_mean_grad$grad_cnt <- as.numeric(unlist(df_county_mean_grad$grad_cnt))
df_county_mean_grad$gradpct <- 100*((df_county_mean_grad$grad_cnt) / (df_county_mean_enroll$enroll_cnt))
df_county_mean_dropout$dropout_cnt <- as.numeric(unlist(df_county_mean_dropout$dropout_cnt))
df_county_mean_dropout$dropoutpct <- 100*((df_county_mean_dropout$dropout_cnt) / (df_county_mean_enroll$enroll_cnt))

df_grad_drop <- data.frame(df_county_mean_grad$county_name ,df_county_mean_grad$gradpct, df_county_mean_dropout$dropoutpct)

df_grad_drop$df_county_mean_grad.gradpct <- round(df_grad_drop$df_county_mean_grad.gradpct, digits=1)
df_grad_drop$df_county_mean_dropout.dropoutpct <- round(df_grad_drop$df_county_mean_dropout.dropoutpct, digits=1)

df_grad_drop$EducationScore = (0.8*df_grad_drop$df_county_mean_grad.gradpct)-(0.2*df_grad_drop$df_county_mean_dropout.dropoutpct) #EducationScore 80% Grad rate - 20% Dropout Rate 

#Ciruclar bar plot code 
df_grad_drop_final <- df_grad_drop[order(-df_grad_drop$df_county_mean_grad.gradpct),] 
df_grad_drop_top10 <- head(df_grad_drop_final, 10)

graph <- plot_ly(
  x = df_grad_drop_top10$df_county_mean_grad.county_name,
  y = df_grad_drop_top10$EducationScore,
  name = "Data",
  type = "bar"
)%>%
  layout(title ="Top 10 Counties for Education",
    xaxis=list(categoryorder="total descending"))
graph

```
Putnam County comes up with a mean graduation percentage of 94% and a mean dropout percentage of 2.9% leading to an Education Score of 74.62. We can also observe from the above plot that the top 10 counties have varying education scores from 75 to 72 indicating that many counties are performing as good as the top county. The graduation percentage of the top counties varies between 94% to 91% whereas the dropout% varies from 5% to 2% leading to an even performance score.


#### **Male v Female enrollment comparison done across the state of New York**

The educational analysis also arises the question of gender disparity within educational institutions. A quick look into the charts gives us a good overview of the recent percentage of male and the female populous in terms of enrolled, graduated, recent degree, recent advanced degree, dropout, still enrolled and local degree students in the state of New York.  




```{r echo=FALSE, warning=FALSE}
#Male v Female Statewide Data comparison (For all Cohorts)

df_male <- df_grad_2021 %>% 
  filter(aggregation_type == 'Statewide') %>% 
  filter(subgroup_name == 'Male') %>% 
 dplyr:: select(enroll_cnt, grad_cnt,local_cnt,  dropout_cnt, reg_cnt, reg_adv_cnt, non_diploma_credential_cnt, still_enr_cnt, ged_cnt)

male_enroll <- sum(df_male$enroll_cnt)
male_grad <- sum(as.integer(df_male$grad_cnt))
male_local <- sum(as.integer(df_male$local_cnt))
male_drop <- sum(as.integer(df_male$dropout_cnt))
male_reg <- sum(as.integer(df_male$reg_cnt))
male_reg_adv <- sum(as.integer(df_male$reg_adv_cnt))
male_stillenrl <- sum(as.integer(df_male$still_enr_cnt))

male_total <- (data.frame(male_enroll, male_grad, male_local, male_drop, male_reg, male_reg_adv, male_stillenrl))


df_female <- df_grad_2021 %>% 
  filter(aggregation_type == 'Statewide') %>% 
  filter(subgroup_name == 'Female') %>% 
 dplyr:: select(enroll_cnt, grad_cnt,local_cnt,  dropout_cnt, reg_cnt, reg_adv_cnt, non_diploma_credential_cnt, still_enr_cnt, ged_cnt)

female_enroll <- sum(df_female$enroll_cnt)
female_grad <- sum(as.integer(df_female$grad_cnt))
female_local <- sum(as.integer(df_female$local_cnt))
female_drop <- sum(as.integer(df_female$dropout_cnt))
female_reg <- sum(as.integer(df_female$reg_cnt))
female_reg_adv <- sum(as.integer(df_female$reg_adv_cnt))
female_nondiploma <- sum(as.integer(df_female$non_diploma_credential_cnt))
female_stillenrl <- sum(as.integer(df_female$still_enr_cnt))
female_ged <- sum(as.integer(df_female$ged_cnt))

female_total <- data.frame(female_enroll, female_grad, female_local, female_drop, female_reg, female_reg_adv, female_nondiploma, female_stillenrl, female_ged)

#Pie chart for Male Population
male_total_graph <- read.csv("Male_df.csv")%>%
drop_na()

fig_4 <- plot_ly(male_total_graph, labels = ~Parameters, values = ~Enroll_cnt, type = 'pie')
fig_4 <- fig_4 %>% layout(title = 'Male data for different parameters across New York')
fig_4
```
Looking at the male population 35.5% are currently enrolled students, 29.7% are graduated students, 16.3% are regent degree holders and 11.9% are regent advance degree holders. Also, a low number is seen in terms of the dropout rate i.e. 2.73%.

```{r echo=FALSE, warning=FALSE}
#Pie chart for female Population
female_total_graph <- read.csv('female_df.csv')%>%
drop_na
fig_5 <- plot_ly(female_total_graph, labels = ~Parameters, values = ~Values, type = 'pie')
fig_5 <- fig_5 %>% layout(title = 'Female data for different parameters across New York')

fig_5
```

 

Similarly with the female populous the number of graduated females is 31.2%, 15.1% are regent degree holders, 14.9% are advanced regent degree holders and only 1.61% are in the dropout count.

The above observations tell us that there is a higher percentage of female graduates and female graduate holding advanced degrees and have a lower dropout count.



### **3.2 Comparative Analysis: Language, Disability, Race, Migrant, Economical status**
 
#### **English vs Non-English Speakers enrollment in different cohorts on a statewide level**

Studies across different cohorts are available giving us an opportunity to investigate the participation across the English-learning population and the non- English learning population in the field of education according to the graduation percentage. 

```{r, echo=FALSE, warning=FALSE}
#English and Non-English Language learners graduation rate
df_english <- df_grad_2021 %>% 
  filter(aggregation_type == 'Statewide') %>% 
  filter(subgroup_name == 'English Language Learner') %>% 
  dplyr::  select(enroll_cnt, grad_cnt, membership_desc) 

df_english$gradpct <- 100*((as.numeric(df_english$grad_cnt)) / (as.numeric(df_english$enroll_cnt)))

df_english_mean <- mean(df_english$gradpct)

df_nonenglish <- df_grad_2021 %>% 
  filter(aggregation_type == 'Statewide') %>% 
  filter(subgroup_name == 'Non-English Language Learner') %>% 
  dplyr::  select(enroll_cnt, grad_cnt, membership_desc) 

df_nonenglish$gradpct <- 100*((as.numeric(df_nonenglish$grad_cnt)) / (as.numeric(df_nonenglish$enroll_cnt)))

df_nonenglish_mean <- mean(df_nonenglish$gradpct)

fig <- df_english %>% plot_ly(labels = ~membership_desc, values = ~gradpct)
fig <- fig %>% add_pie(hole = 0.7)
fig <- fig %>% layout(title = "English Learners Average graduation percentage by different co-horts",  showlegend = T,
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
fig

fig1 <- df_nonenglish %>% plot_ly(labels = ~membership_desc, values = ~gradpct)
fig1 <- fig1 %>% add_pie(hole = 0.7)
fig1 <- fig1 %>% layout(title = " Non-English Learners Average graduation percentage by different co-horts",  showlegend = T,
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
fig1


```

The wheel shows the percentage count of the English-learning population across different cohorts. Amongst the English-learning population the 2017 cohort- 4year outcome, indicated by the color ‘blue’ has the maximum graduation percentage as shown in the chart. 

The wheel shows the percentage count of the Non-English learning population across different cohorts. Amongst the Non-English learning population, the 2016 cohort- 5-year outcome and 2015 cohort – 6-year outcome has the highest graduation percentage amongst others. 

From the above, we can conclude that the overall graduation rate among English and Non-English learners in the field of education for the state has only slight variation.  


#### **Parameter comparison of Disabled students and All students**


The education system has been supportive of the disabled community helping them by providing the necessary arrangements in times of need. This assumption is put to test via the analysis conducted.


```{r Disability and all students, echo=FALSE, warning=FALSE}
#disability va all student








df <- read.csv("GRAD_RATE_AND_OUTCOMES_2021.csv")





df_disab <- df %>%
   filter(df$aggregation_type=='County') %>%
 dplyr::  select(subgroup_name,grad_pct, dropout_pct, county_name) %>%
   filter(subgroup_name == 'Students with Disabilities') %>%
   group_by(county_name)





df_allstd <- df %>%
   filter(df$aggregation_type=='County') %>%
dplyr::   select(subgroup_name,grad_pct, dropout_pct, county_name) %>%
   filter(subgroup_name == 'All Students') %>%
   group_by(county_name)





df_disab$grad_pct <- gsub('%',"",df_disab$grad_pct)
df_disab$grad_pct <- as.numeric(df_disab$grad_pct)




df_disab$dropout_pct <- gsub('%',"",df_disab$dropout_pct)
df_disab$dropout_pct <- as.numeric(df_disab$dropout_pct)




df_allstd$grad_pct <- gsub('%',"",df_allstd$grad_pct)
df_allstd$grad_pct <- as.numeric(df_allstd$grad_pct)




df_allstd$dropout_pct <- gsub('%',"",df_allstd$dropout_pct)
df_allstd$dropout_pct <- as.numeric(df_allstd$dropout_pct)



df_disab_county <- df_disab %>%
  dplyr:: select(county_name, grad_pct,dropout_pct) %>%
   group_by(county_name) %>%
   summarise_at(c('grad_pct', 'dropout_pct'), ~round(mean(.),0), na.rm = TRUE) %>%
   arrange(desc(grad_pct)) %>%
   slice(1:10)



df_allstd_county <- df_allstd %>%
dplyr::   select(county_name, grad_pct,dropout_pct) %>%
   group_by(county_name) %>%
   summarise_at(c('grad_pct', 'dropout_pct'), ~round(mean(.),0), na.rm = TRUE) %>%
   arrange(desc(grad_pct)) %>%
   slice(1:10)
   
df_disab_county$county_name <-factor(df_disab_county$county_name, levels = c("PUTNAM", "NASSAU","SUFFOLK","ROCKLAND", "GENESEE", "WESTCHESTER", "WYOMING", "RENSSELAER", "ESSEX", "SCHOHARIE"
))



df_allstd_county$county_name <-factor(df_allstd_county$county_name, levels = c("NASSAU", "PUTNAM","GENESEE","ESSEX", "HAMILTON", "LIVINGSTON", "ONTARIO", "SARATOGA", "SCHUYLER", "WESTCHESTER"
))




ggplot(df_disab_county,aes(x=county_name, y=grad_pct,fill=county_name)) +
  geom_bar(stat="identity", color="black", position=position_dodge(),width=0.75,size=0.3)+labs(x="Counties",y="Graduation Rate",title="County-wise Graduation rate for people with Disabilities")+theme(plot.title = element_text(hjust=0.5))+guides(fill=guide_legend(title="Counties"))+

 
  coord_polar(theta = "x",start=0) +
  scale_fill_brewer(palette="spectral")+
  theme_light()+
  theme( axis.text.y = element_text(size = 12,colour="black"),
         axis.text.x=element_text(size = 7,colour="black"))



ggplot(df_allstd_county,aes(x=county_name, y=grad_pct,fill=county_name)) +
  geom_bar(stat="identity", color="black", position=position_dodge(),width=0.75,size=0.3)+labs(x="Counties",y="Graduation Rate",title="County-wise Graduation rate for All Students")+theme(plot.title = element_text(hjust=0.5))+
  coord_polar(theta = "x",start=0) +guides(fill=guide_legend(title="Counties"))+
  scale_fill_brewer(palette=7)+
  theme_light()+
  theme( axis.text.y = element_text(size = 12,colour="black"),
         axis.text.x=element_text(size = 7,colour="black"))

```
The above chart shows the graduation percentage across various counties of all students in the state of New York. The counties of Nassau County and Putnam County leads with a graduation percentage of 94% and is followed by, Genesee County(93%) among others as shown above having a varying graduation rate of 94% to 92%  

Whereas the assumption might not be true. Although Putnam County comes in lead with the previous observation having the graduation percentage of 85% but falls overall with a low in Schoharie County the graduation percentage of 73% and an overall variation of 85% to 73%.

Although the system seems supportive of providing equal opportunities in the field of education further actions should be taken to improve the conditions in counties having low graduation percentage.


#### **Migrant enrollments for the year 2020 and 2021 by County** 
A study across the migrated Populous in the year 2020 vs the migrated Populous in the year 2021 is conducted to find and compare the participation level of both communities.
```{r Migrant graduation comparison for 2 years, echo=FALSE, warning=FALSE}
#Migrant vs Non-migrant Population in New York State and their Graduate percentage

df_migrant <- df_grad_2021 %>% 
 dplyr:: select(subgroup_name, aggregation_type, county_name, enroll_cnt, grad_pct) %>% 
  filter(subgroup_name == 'Migrant') %>%
  filter(aggregation_type=='County') %>%
  group_by(county_name) %>% 
  na.omit()

migrant_enrol <- aggregate(enroll_cnt ~ county_name, data=df_migrant, sum)
migrant_enrol_desc <- head(migrant_enrol[order(-migrant_enrol$enroll_cnt),],7)

df_migrant_2020 <- df_grad_2020 %>% 
dplyr::  select(subgroup_name, aggregation_type, county_name, enroll_cnt, grad_pct) %>% 
  filter(subgroup_name == 'Migrant') %>%
  filter(aggregation_type=='County') %>%
  group_by(county_name) %>% 
  na.omit()

migrant_enrol_2020 <- aggregate(enroll_cnt ~ county_name, data=df_migrant_2020, sum)
migrant_enrol_desc_2020 <- head(migrant_enrol_2020[order(-migrant_enrol_2020$enroll_cnt),],7)

df_2021_2020 <- data.frame(migrant_enrol_desc, migrant_enrol_desc_2020)

poi <- plot_ly(df_2021_2020, x = ~county_name, y = ~enroll_cnt, type = 'bar', name = 'Migrant_2021 ',
        marker = list(color = 'rgb(55, 83, 109)'))
poi <- poi %>% add_trace(y = ~enroll_cnt.1, name = 'Migrant_2020', marker = list(color = 'rgb(26, 118, 255)'))
poi <- poi %>% layout(title = 'Migrant Enrollments:2020 vs 2021',
         xaxis = list(
           title = "Counties",
           tickfont = list(
             size = 14,
             color = 'rgb(107, 107, 107)')),
         yaxis = list(
           title = 'Enrollments',
           titlefont = list(
             size = 16,
             color = 'rgb(107, 107, 107)'),
           tickfont = list(
             size = 14,
             color = 'rgb(107, 107, 107)')),
         legend = list(x = 0, y = 1, bgcolor = 'rgba(255, 255, 255, 0)', bordercolor = 'rgba(255, 255, 255, 0)'),
         barmode = 'group', bargap = 0.15, bargroupgap = 0.1)

poi  

```

The above plot showcases the difference in enrollment numbers in Chautauqua County, Columbia County, Franklin County, Orange County, Saint Lawrence County, Suffolk County, and Sullivan County, in terms of enrollment count.

It can be observed that the enrollment counts his doing almost similar in some of the counties over the period of 2020 to 2021 whereas there has been a slight drop in numbers of Chautauqua County and Suffolk County.


#### **Race Wise comparison of Dropout Rate and Graduation Rate**

To get an overview of the education parameter of different racial segments of the population to understand the understand how to better help them.


```{r echo=FALSE, warning=FALSE}
library(plotly)

df_race <- read.csv("GRAD_RATE_AND_OUTCOMES_2021.csv") %>% 
filter(aggregation_type == "County") %>% 
dplyr::select(subgroup_name, grad_pct,dropout_pct) %>%
 filter(subgroup_name == "American Indian or Alaska Native" | subgroup_name == 'Asian or Native Hawaiian/Other Pacific Islander'|subgroup_name == 'Black or African American'|subgroup_name == 'Hispanic or Latino'|subgroup_name == 'White'|subgroup_name == 'Multiracial') 



df_race$grad_pct<-gsub("%", " ",df_race$grad_pct)
df_race$grad_pct <-gsub("-", "0",df_race$grad_pct)
df_race$grad_pct <- as.numeric(df_race$grad_pct)




df_race$dropout_pct<- gsub("%", " ",df_race$dropout_pct)
df_race$dropout_pct <- gsub("-", "0",df_race$dropout_pct)
df_race$dropout_pct <- as.numeric(df_race$dropout_pct)

df_race5 <- df_race %>%  
  dplyr::select(subgroup_name, grad_pct,dropout_pct) %>%
   group_by(subgroup_name) %>%
   summarise_at(c('grad_pct', 'dropout_pct'), mean, na.rm = TRUE)

fig <- plot_ly(df_race5, x = ~subgroup_name, y = ~grad_pct, type = 'bar', name = 'Graduation Rate')
fig <- fig %>% add_trace(y = df_race5$dropout_pct, name = 'Dropout Rate')
fig <- fig %>% layout(xaxis=list(title='Races'), yaxis = list(title = 'Percentage'), barmode = 'stack')

fig

```
The analysis shows how the graduation percentage and dropout percentage vary across the State of New York with the Caucasian population having the highest graduation percentage of 87.47% followed by the Hispanic Populus(70.5%) and the African American Populus(67.4%).

It is also observed that the dropout percentage is led by the Hispanic Population with value of 10.3% followed by the African American Populus(8.6%), and multiracial Populus(6.7%) among others.


#### **Comparison of graduation and dropout rate for Economically v Non-Economically challenged sections of the population** 
Economic Analysis is necessary for the revision of budget allocation among different counties, a comparative analysis is conducted considering the graduation percentage and dropout percentage between economically and non-Economically disadvantaged groups.



```{r echo=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(plotly)





df <- read.csv("GRAD_RATE_AND_OUTCOMES_2021.csv")





df_eco <- df %>%
   filter(df$aggregation_type=='County') %>%
dplyr::   select(subgroup_name,grad_pct, dropout_pct, county_name) %>%
   filter(subgroup_name == 'Economically Disadvantaged') %>%
   group_by(county_name)





df_neco <- df %>%
   filter(df$aggregation_type=='County') %>%
  dplyr:: select(subgroup_name,grad_pct, dropout_pct, county_name) %>%
   filter(subgroup_name == 'Not Economically Disadvantaged') %>%
   group_by(county_name)





df_eco$grad_pct <- gsub('%',"",df_eco$grad_pct)
df_eco$grad_pct <- as.numeric(df_eco$grad_pct)




df_eco$dropout_pct <- gsub('%',"",df_disab$dropout_pct)
df_eco$dropout_pct <- as.numeric(df_eco$dropout_pct)




df_neco$grad_pct <- gsub('%',"",df_allstd$grad_pct)
df_neco$grad_pct <- as.numeric(df_neco$grad_pct)




df_neco$dropout_pct <- gsub('%',"",df_neco$dropout_pct)
df_neco$dropout_pct <- as.numeric(df_neco$dropout_pct)



df_eco_county <- df_eco %>%
  dplyr:: select(county_name, grad_pct,dropout_pct) %>%
   group_by(county_name) %>%
   summarise_at(c('grad_pct', 'dropout_pct'), ~round(mean(.),0), na.rm = TRUE) %>%
   arrange(desc(grad_pct)) %>%
   slice(1:10)



df_neco_county <- df_neco %>%
 dplyr::  select(county_name, grad_pct,dropout_pct) %>%
   group_by(county_name) %>%
   summarise_at(c('grad_pct', 'dropout_pct'), ~round(mean(.),0), na.rm = TRUE) %>%
   arrange(desc(grad_pct)) %>%
   slice(1:10)



merge_eco_neco<- merge(df_eco_county,df_neco_county,by='county_name') %>% drop_na()



#plotly grouped bar plot_grad
men_econeco_grad <- plot_ly(merge_eco_neco, x = ~county_name, y = ~grad_pct.x, type = 'bar', name = 'Economically Disadvantaged ',
        marker = list(color = 'rgb(55, 83, 109)'))
men_econeco_grad <- men_econeco_grad %>% add_trace(y = ~grad_pct.y, name = 'Not Economically Disadvantaged', marker = list(color = 'rgb(26, 118, 255)'))
men_econeco_grad <- men_econeco_grad %>% layout(title = 'Grad. Percentage For Economically V Non Economically',
         xaxis = list(
           title = "",
           tickfont = list(
             size = 14,
             color = 'rgb(107, 107, 107)')),
         yaxis = list(
           title = 'Grad. %',
           titlefont = list(
             size = 16,
             color = 'rgb(107, 107, 107)'),
           tickfont = list(
             size = 14,
             color = 'rgb(107, 107, 107)')),
         legend = list(x = 0, y = 1, bgcolor = 'rgba(255, 255, 255, 0)', bordercolor = 'rgba(255, 255, 255, 0)'),
         barmode = 'group', bargap = 0.15, bargroupgap = 0.1) %>%
         layout(legend = list(x = 0.7, y = 1.05))



men_econeco_grad




#plotly grouped bar plot_drop
men_econeco_drop <- plot_ly(merge_eco_neco, x = ~county_name, y = ~dropout_pct.x, type = 'bar', name = 'Economically Disadvantaged ',
        marker = list(color = 'rgb(55, 83, 109)'))
men_econeco_drop <- men_econeco_drop %>% add_trace(y = ~dropout_pct.y, name = 'Not Economically Disadvantaged', marker = list(color = 'rgb(26, 118, 255)'))
men_econeco_drop <- men_econeco_drop %>% layout(title = 'Dropout Percentage For Economically V Non Economically',
         xaxis = list(
           title = "",
           tickfont = list(
             size = 14,
             color = 'rgb(107, 107, 107)')),
         yaxis = list(
           title = 'Droupout Percentage',
           titlefont = list(
             size = 16,
             color = 'rgb(107, 107, 107)'),
           tickfont = list(
             size = 14,
             color = 'rgb(107, 107, 107)')),
         legend = list(x = 0, y = 1, bgcolor = 'rgba(255, 255, 255, 0)', bordercolor = 'rgba(255, 255, 255, 0)'),
         barmode = 'group', bargap = 0.15, bargroupgap = 0.1) %>%
         layout(legend = list(x = 0.7, y = 1.05))



men_econeco_drop
```
An even graduation percentage is seen for both the groups with the highest graduation rate in Genesee County for the economically challenged community(90%) and a variation from 90% to 86%; and Nassau and Putnam having similar high graduation rates for the non-economically challenged community(90%).

A wide disparity is seen in the dropout rate between the two communities with the economically challenged community having a variation of 14% to 2%, Essex County having the highest dropout percentage and the non-economically challenged community having a variation of 3% to 1%.
We can conclude that a revised budget for the affected counties can be helpful in providing support in terms of education in affected areas.

### **4.  Conclusion**

The above analysis helps us observe the condition across various territories included through which we reach the conclusions as follows,

1.	The counties having the highest population are among the counties also having the highest enrollment rate as assumed and bolstered by the Enrollment Comparison by County analysis and School Comparison by highest enrollment analysis respectively.

2.	Good standardization is observed within the system as the counties show a close number in terms of education scores. 

3.	The female populous has higher percentage numbers in graduates and graduates holding advanced degrees with low dropout rates whereas the male populous is also doing well in terms of graduation and dropout percentage.

4.	The overall participation among English and non-English speakers in the field of education is almost similar supporting the effectiveness of the system for people of different languages.

5.	The system seems supportive of providing equal opportunities in the field of education, but further actions should be taken to support the diverse population while considering the African American and Hispanic populous.

6.	The education system is also doing almost similar in many counties over the time of 2020 to 2021 with a slight rise in numbers in Chautauqua County and Suffolk County for the migrants supporting the effective systems in those areas.

7.	Further analysis should be done to find the cause of the high dropout rate among multiracial communities while, proper funds allocation to ensure equal opportunities despite the economic gap.


### **5.References**

•	https://www.newyork-demographics.com/counties_by_population

•	https://www.datanovia.com/

•	https://r-graph-gallery.com/

• https://www.data-to-viz.com/

•	https://plotly.com/r/

•	https://www.statology.org/r-plot-change-axis-scale/ 

•	https://data.nysed.gov/downloads.php 

•	https://www.datamentor.io/r-programming/histogram/

•	https://community.plotly.com/t/title-alignment-python/30820

•	https://plotly.com/r/bar-charts/
























